\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-11-09 abstract.tex
address comment in abstract
took a pass on §1–3

With the growing practice of mechanizing language metatheories,
it has become ever more pressing that interactive theorem provers 
make it easy to write reusable, extensible code and proofs.
%
This paper presents a language design geared towards extensible metatheory
mechanization in a proof assistant.
The new design achieves reuse and extensibility via a form of family
polymorphism, a seemingly object-oriented idea, that allows code and
proofs to be polymorphic to their enclosing families.
Our development addresses significant technical challenges that arise
from the underlying language of a proof assistant being simultaneously
functional, dependently typed, a logic, and an interactive tool.
%
Our results include
\begin{enumerate*}
\item a prototypical implementation of the language design as a Coq plugin,
\item a formalization of the language design as a constructive type theory with inductive types,
\item proofs of strong guarantees\EDJ{We don't yet have normalization though} afforded by the type theory,
and 
\item case studies showing how the new expressiveness naturally addresses real
programming challenges in metatheory mechanization.
\end{enumerate*}

>>>>

With the growing practice of mechanizing language metatheories,
it has become ever more pressing that interactive theorem provers 
make it easy to write reusable, extensible code and proofs.
%
This paper presents a language design geared towards extensible metatheory
mechanization in a proof assistant.
The new design achieves reuse and extensibility via a form of family
polymorphism, a seemingly object-oriented idea, that allows code and
proofs to be polymorphic to their enclosing families.
Our development addresses technical challenges that arise
from the underlying language of a proof assistant being simultaneously
functional, dependently typed, a logic, and an interactive tool.
%
Our results include
\begin{enumerate*}
\item a prototypical implementation of the language design as a Coq plugin,
\item a formalization of the language design as a constructive type theory with inductive types,
\item consistency and canonicity of the type theory,
%\EDJ{We don't yet have normalization though}
and 
\item case studies showing how the new expressiveness naturally addresses real
programming challenges in metatheory mechanization.
\end{enumerate*}

### 2022-11-09 challenge.tex
AI case study
took a pass on §1–3

\noindentparagraph{%
\headinglabel{C4}{chg:tooling}.\,
User experience and tooling.%
}

Interactive theorem proving and tactic programming are central to
the idiomatic programming experience in a proof assistant.
A language design integrating family polymorphism should be compatible
with these programming styles.
In particular, it should be possible in our tool to
incrementally navigate through vernacular commands and, moreover, construct proofs
with common tactics while getting instant feedback on the proof state,
even in the middle of a family definition.
In addition, it should remain possible to normalize terms that mention families.

>>>>

\noindentparagraph{%
\headinglabel{C4}{chg:tooling}.\,
User experience and system implementation.%
}

Interactive theorem proving and tactic programming are central to
a typical programming experience with a proof assistant.
A language design integrating family polymorphism should be compatible
with these forms of programming.
In particular, it should be possible in our system to
incrementally navigate through vernacular commands and, moreover, construct proofs
with common tactics while getting instant feedback on the proof state,
even in the middle of a family definition.
Last but not least, in addition to producing theorems,
it should be possible for terms defined in families to possess computational content.

### 2022-11-09 introduction.tex
add one sentence about parser case study
took a pass on §1–3

In the Coq proof assistant~\cite{coq}, for instance, inductive types, as well as functions and
theorems over inductive types, are closed to extension.
So to reuse mechanized metatheories,
the common practice is still to copy code and proofs and then modify them in each extension.
But having to maintain multiple copies is highly non-modular and
antithetical to good software engineering.
%
The programmer could also turn to design patterns~\cite{delaware2011,delaware2013} that
use software product lines or Church encodings.
But they tend to require heavy lifting from the programmer to make code
extensible, often leading to non-idiomatic programming styles.

@@

\paragraph{Contributions}

@@

\begin{itemize}[leftmargin=3.5ex]

\item We present a
%\EDJ{Are you sure? This claim is so bold.}\YZreply{Is there any previous language designs in this space?}\EDJreply{Olivier Boite. 2004. Proof reuse with extended inductive types. In International Conference on Theorem Proving in
%Higher Order Logics. Springer, 50–65.  I think I also mentioned several paper in the related work.}
language design that enables extensible
metatheory mechanization in a higher-order, dependent type theory with
inductive types (\cref{sec:lang-design}).
The language design reconciles the expressiveness enabled by
family polymorphism with the rigor of a proof assistant,
while retaining an idiomatic programming style.

@@

\item We present case studies of using our Coq plugin to mechanize
language metatheories (\cref{sec:casestudies}).
They show how our language design naturally solves the EP and
enables a high degree of reuse and extensibility
for proof engineering.

>>>>

In the Coq proof assistant~\cite{coq}, for instance, inductive types, as well as functions and
theorems over inductive types, are closed to extension.
So to reuse mechanized metatheories,
the common practice is still to copy code and proofs and then modify them in each extension.
But having to maintain multiple copies is highly non-modular and
antithetical to good software engineering.
%
The programmer could also turn to design patterns and encodings~\cite{delaware2011,delaware2013,schwaab2013modular,keuchel2013generic,forsta2020}.
But they tend to require heavy lifting from the programmer to make code
extensible and often lead to non-idiomatic programming styles.

@@

\noindentparagraph{Contributions}

@@

\begin{itemize}%[leftmargin=3.5ex]
    [labelsep=*,leftmargin=1pc,itemsep=3pt]

\item We present a
%\EDJ{Are you sure? This claim is so bold.}\YZreply{Is there any previous language designs in this space?}\EDJreply{Olivier Boite. 2004. Proof reuse with extended inductive types. In International Conference on Theorem Proving in
%Higher Order Logics. Springer, 50–65.  I think I also mentioned several paper in the related work.}
language design that enables extensible
metatheory mechanization in a higher-order, dependent type theory with
inductive types (\cref{sec:lang-design}).
The language design reconciles the expressiveness enabled by
family polymorphism with the rigor of a proof assistant,
while largely retaining an idiomatic programming style.

@@

\item We present case studies of using our Coq plugin to mechanize
language metatheories (\cref{sec:casestudies}).
They show how our language design naturally solves the EP and
enables a good amount of reuse and extensibility
for mechanizing proofs.
